### 2021-04-27 scripts/EuringPosterFigures.Rmd
bias across life history scenarios

---
title: "EuringPosterFigures"
author: ""
date: "4/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, results = 'asis', fig.width = 4, fig.height = 3.5)

library(tidyr)
library(dplyr)
library(cowplot)
library(ggplot2)
library(coda)
library(captioner)
library(knitr)
library(reshape2)
library(here)

```

```{r data load}

load(file = here::here('results', 'EuringPoster', 'processedIPMOutput.RData'))

row.low <- do.call(rbind, lapply( ls(patt='lowout'), get) )
row.med <- do.call(rbind, lapply( ls(patt='medout'), get) )
row.high <- do.call(rbind, lapply( ls(patt='highout'), get) )
rm(list=grep('highout|medout|lowout',ls(),value=TRUE,invert=FALSE))

low <- row.low %>%
  dplyr::rename(phi1 = `mean.phi[1]`, phiad = `mean.phi[2]`)

low.params <- readRDS(file = here::here('data', 'low.lam.params.rds')) %>%
  transform(scenario = 1:25)

med <- row.med %>%
  dplyr::rename(phi1 = `mean.phi[1]`, phiad = `mean.phi[2]`)

med.params <- readRDS(file = here::here('data', 'med.lam.params.rds')) %>%
  transform(scenario = 1:25)

high <- row.high %>%
  dplyr::rename(phi1 = `mean.phi[1]`, phiad = `mean.phi[2]`)

high.params <- readRDS(file = here::here('data', 'high.lam.params.rds')) %>%
  transform(scenario = 1:25)

#combine to get true values integrated in results
low.dat <- low %>%
  inner_join(low.params, by = 'scenario', suffix = c('.obs', '.true')) %>%
  transform(lambda.scenario = 'Decreasing')

med.dat <- med %>%
  inner_join(med.params, by = 'scenario', suffix = c('.obs', '.true')) %>%
  transform(lambda.scenario = 'Stable')

high.dat <- high %>%
  inner_join(high.params, by = 'scenario', suffix = c('.obs', '.true')) %>%
  transform(lambda.scenario = 'Increasing')

all.dat <- bind_rows(low.dat, med.dat, high.dat)

```

```{r bias}

#get posterior medians
all.meds <- all.dat %>%
  dplyr::select('phi1.obs', 'phi1.true', 'phiad.obs', 'phiad.true', 'fec.obs', 'fec.true',
                            'sims', 'scenario', 'simscenarios', 'lambda.scenario') %>%
  reshape2::melt(id.vars = c('lambda.scenario', 'sims', 'scenario', 'simscenarios')) %>%
  group_by(lambda.scenario, scenario, simscenarios, variable) %>%
  summarize(median = median(value), .groups = 'keep') %>%
  reshape2::dcast(lambda.scenario + scenario + simscenarios ~ variable, value.var = 'median')

#just bias
rel.bias <- all.meds %>%
  #calculate relative bias, mean across scenarios
  transform(phi1.bias = (phi1.obs-phi1.true)/phi1.true,
            phiad.bias = (phiad.obs-phiad.true)/phiad.true,
            fec.bias = (fec.obs-fec.true)/fec.true) %>%
  dplyr::select('lambda.scenario', 'scenario', 'simscenarios', 
                'phi1.bias', 'phiad.bias', 'fec.bias') %>%
  reshape2::melt(id.vars = c('lambda.scenario', 'scenario', 'simscenarios')) %>%
  group_by(lambda.scenario, simscenarios, variable) %>%
  dplyr::summarize(mean.rel.bias = mean(value), .groups = 'keep') %>%
  transform(simscenarios = factor(simscenarios, levels = c(1:9), 
                                  labels = rep(c('L', 'M', 'H'), 3)))

ggplot(rel.bias, aes(x = factor(simscenarios), y = variable, fill = mean.rel.bias)) +
  geom_tile() +
  xlab('Detection level') + ylab('') +
  facet_grid(~lambda.scenario, drop = T, scales = 'free_x') +
  scale_fill_gradient2(name = "Relative bias",
                      mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
  theme_bw() +
  theme(legend.position = 'top',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"))

```

```{r rmse}

#for the love of buddha, someone please check this calculation of rmse
rmse.vals <- all.meds %>%
  transform(rmse.fec = (fec.obs-fec.true)^2,
            rmse.phiad = (phiad.obs-phiad.true)^2,
            rmse.phi1 = (phi1.obs-phi1.true)^2) %>%
  dplyr::select(lambda.scenario, scenario, simscenarios, rmse.fec, rmse.phiad, rmse.phi1) %>%
  reshape2::melt(id.vars = c('lambda.scenario', 'scenario', 'simscenarios')) %>%
  group_by(simscenarios, lambda.scenario, variable) %>%
  dplyr::summarize(rmse = sqrt(mean(value)), .groups = 'keep') %>%
  transform(simscenarios = factor(simscenarios, levels = c(1:9), 
                                  labels = rep(c('L', 'M', 'H'), 3)))

#dunno about this - looks suspiciously similar across detection levels :(
ggplot(rmse.vals, aes(x = factor(simscenarios), y = variable, fill = rmse)) +
  geom_tile() +
  xlab('Detection level') + ylab('') +
  facet_grid(~lambda.scenario, drop = T, scales = 'free_x') +
  scale_fill_gradient2(name = "RMSE",
                      mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
  theme_bw() +
  theme(legend.position = 'top',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"))

```

```{r bias across life history scenarios}

##Same as above but looking at the variability across data-generating true values for phiad, phi1, fec

#get posterior medians; same as above except keep 'scenario' in group_by
all.meds <- all.dat %>%
  dplyr::select('phi1.obs', 'phi1.true', 'phiad.obs', 'phiad.true', 'fec.obs', 'fec.true',
                            'sims', 'scenario', 'simscenarios', 'lambda.scenario') %>%
  reshape2::melt(id.vars = c('sims', 'scenario', 'simscenarios', 'lambda.scenario')) %>%
  group_by(lambda.scenario, scenario, simscenarios, variable) %>%
  summarize(median = median(value)) %>%
  reshape2::dcast(lambda.scenario + scenario + simscenarios ~ variable, value.var = 'median') %>%
  #calculate relative bias
  transform(phi1.bias = (phi1.obs-phi1.true)/phi1.true,
            phiad.bias = (phiad.obs-phiad.true)/phiad.true,
            fec.bias = (fec.obs-fec.true)/fec.true)

#just bias; same as above, keep 'scenario'
rel.bias <- all.meds %>%
  dplyr::select('lambda.scenario', 'scenario', 'simscenarios', 
                'phi1.bias', 'phiad.bias', 'fec.bias') %>%
  reshape2::melt(id.vars = c('lambda.scenario', 'scenario', 'simscenarios')) %>%
  transform(simscenarios = factor(simscenarios, labels = rep(c('L', 'M', 'H'),3)))

#grab true values from above
true_vals <- all.dat %>%
  distinct(lambda.scenario, scenario, phiad.true, fec.true, phi1.true) %>%
  transform(phiad.true = round(phiad.true,1),
            phi1.true = round(phi1.true,1), fec.true = round(fec.true,1))

#merge back onto bias df so can have something to name/view the 'scenarios' 
rel.bias.vals <- rel.bias %>%
  merge(true_vals, by = c('lambda.scenario', 'scenario'))

#scenario number -- not a useful view, but inspired this investigation
ggplot(rel.bias, aes(x = factor(scenario), y = variable, fill = value)) +
  geom_tile() +
  xlab('Scenario') + ylab('Parameter') +
  facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
  scale_fill_gradient2(name = "Relative bias",
                      mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
  theme_bw() +
  theme(legend.position = 'top',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"))

#view by fec.true
ggplot(rel.bias.vals, aes(x = factor(fec.true), y = variable, fill = value)) +
  geom_tile() +
  xlab('True fecundity') + ylab('Parameter') +
  facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
  scale_fill_gradient2(name = "Relative bias",
                      mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
  theme_bw() +
  theme(legend.position = 'top',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
        axis.text.x = element_text(angle = 90, vjust = 0.5))

#view by phiad.true
ggplot(rel.bias.vals, aes(x = factor(phiad.true), y = variable, fill = value)) +
  geom_tile() +
  xlab('True phi.ad') + ylab('Parameter') +
  facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
  scale_fill_gradient2(name = "Relative bias",
                      mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
  theme_bw() +
  theme(legend.position = 'top',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
        axis.text.x = element_text(angle = 90, vjust = 0.5))

#view by phi1.true
ggplot(rel.bias.vals, aes(x = factor(phi1.true), y = variable, fill = value)) +
  geom_tile() +
  xlab('True phi.1') + ylab('Parameter') +
  facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
  scale_fill_gradient2(name = "Relative bias",
                      mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
  theme_bw() +
  theme(legend.position = 'top',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
        axis.text.x = element_text(angle = 90, vjust = 0.5))

```