### 2021-08-17 README.Rmd
repo reorganization

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Understanding the processes that control population dynamics in wild populations is a critical component of conservation and management but is frequently limited by missing demographic information, uncertainty, and temporally or spatially misaligned datasets (Schaub & Abadi 2011, Zipkin & Saunders 2017). Integrated population models (IPMs; Besbeas et al. 2002, Brooks et al. 2004) are an increasingly popular tool in ecology that can help overcome these limitations by combining disparate datasets in a single, unified analysis (Schaub & Abadi 2011). This approach can provide several advantages, such as improved precision (Schaub et al. 2007, Tavecchia 2009, Abadi et al. 2010) and the estimation of parameters that would otherwise be unidentifiable (Besbeas 2005, Schaub et al. 2007, Veran & Lebreton 2008), which is often the case for key demographic rates such as recruitment or immigration. These advances have practical benefits for resource managers that must make decisions based on limited available information. As IPMs become more widely used, however, it is critical that they are robustly evaluated. In particular, the limitations of the IPM framework are still underexplored, particularly in situations where available data are sparse, of poor quality, or when data to inform specific parameters are entirely absent.

Through a simulation analysis, we will identify monitoring programs and demographic scenarios where integrated population models may not lead to improved precision or return unbiased estimates of parameters that would otherwise be unidentifiable when certain datasets are excluded. We apply this simulation study to to a passerine bird life history. We assume all model assumptions are met and compare parameter estimation for a traditional IPM with data informing abundance (count data), survival (mark-recapture data), and reproductive output (nest monitoring data) to situations when either of the latter two datasets are omitted. In doing so, we will refine our understanding of the mechanisms underlying IPMs and can better identify when they may serve as a useful tool or might not be necessary or warrant the collection of additional data. Improving our understanding of the strengths and weaknesses of the IPM framework can provide helpful insights for the design and implementation of field survey programs so as to most efficiently make use of limited resources to inform management efforts aimed at monitoring or recovering species that are deemed ecosystem indicators or depleted and in need of conservation intervention.

# Scripts, current version

## 0 - preparing scenarios

### compute_time_calc.R
Code for approximating the total run time, given number of simulated datasets, number of parameter scenarios, computer cores, and estimated run time for each model.

### generate_scenarios.R

Function 'getNviable' for finding the combinations of parameters that give a population growth rate ($\lambda$) within certain bounds, using the eigenvalue from the Leslie matrix. The output from this function produces the parameter scenarios used to simulate population trajectories, observation data from the trajectories, and fit the the IPM models.

## 1 - simulating data

### IPM_sim_2.0function.R

Script contains two functions for simulating data: 'simPopTrajectory' and 'simData'.

The 'simPopTrajectory' function simulates the true population trajectory by individual and age class, starting from the stable age distribution. Individual fates are retained throughout total number of years.

The 'simData' function uses the output from the 'simPopTrajectory' function to simulate the observation process for each data type. The true individual data is broken down into independent sets, each is subject to observation error. Mark-resight, count, and reproductive success data are output from this function.

### simulateTrajectories.R

Script that pulls from the scenarios that were created in '0 - preparing scenarios' and the functions in IPM_sim_2.0function.R to simulate observation data sets for model fitting.

## 2 - models

### IPMinitvalues.R
Functions for putting data in the form required by the survival model.

### IPM_marray.R
Script contains 4 IPM NIMBLE models.

 1. full IPM model ('IPMmod') where all three datasets are included
 2. model without productivity dataset ('nonests')
 3. model without survival model ('nomr')
 4. abundance only model without productivity or survival ('abundonly')

## 3 - run models

### run_scenarios_helperFns.R

Script with functions and code to run NIMBLE models. Function 'marray' transforms the survival data capture histories to m-array format. MCMC settings in this script, with functions for each of the 4 models. Each model has a run model function that defines constants, initial values, data, parameters to monitor, and code to build/run the models in NIMBLE.

### run_scenarios.R

Main workhorse script that pulls from the scripts above to run models in parallel and output results.

## 4 - process results

### 01_load_and_process_data.R

Script that checks the Gelman-Rubin convergence statistic for each output MCMC sample, then thins the chains to reduce file size and places the output into categories based on the population growth rate, $\lambda$.

### 02_compute_geom_means.R

Computes the geometric means for $\lambda$ in each MCMC output.

# Data

Contains true parameter scenarios used to simulate population trajectories and observation data, all viable parameter combinations in files '-.lam.combos.RDS' and the selected parameter combinations used to simulate data are in files '-.lam.params.RDS' *(where "-" is low, med, or high)*. Folders labeled '-Trajectories' contain the simulated population trajectories for each of the scenarios found in '-.lam.params.RDS'. Each simulated trajectory is labeled with the scenario and the simulation number for that scenario. For example, in the 'highTrajectory' folder: 'highpopTraj-1-2.RDS' corresponds to the first parameter scenario in 'high.lam.param.RDS' and is the second population trajectory simulated using those parameter values.

File 'true.vals.csv' is the true parameter values used in each scenario. 'scenario_ID.csv' tracks the combinations of detection probabilities and IPM models.