### 2024-10-02 README.Rmd
Update abstract; add title and authors

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Effects of species life history and data availability on integrate population model performance

Abby E. Bratt, Caroline D. Cappello, Amelia J. DuVall, Hannah A. Sipe, Amanda J.
Warlick, Beth Gardner, Sarah J. Converse

Code provided for peer review and may change

## Abstract

Integrated population models (IPMs) are an increasingly popular tool for population modeling in ecology. By combining datasets in a unified analysis, IPMs link demographic rates with population dynamics and can thereby reduce bias, improve precision, and estimate parameters that would otherwise be unidentifiable using traditional approaches. However, the performance of IPMs in different situations remains underexplored and the conditions under which IPMs produce more precise estimates compared with those produced in singular analyses are poorly understood.

We assessed the performance of IPMs across a range of life history parameters, population trajectories, and data availability scenarios using simulated count, mark-resight, and productivity data typical of a passerine monitoring program. We examined relative bias and root mean squared error for abundance, trend in abundance, survival, and fecundity across models when all datasets were included versus when one or two datasets were omitted or when the included datasets were of poorer quality (i.e., lower detection probabilities).

Parameter estimates generally showed higher bias when the datasets that more directly informed those parameters were omitted or when the true data-generating parameters were representative of more extreme life history traits (e.g., high survival and low reproduction that characterizes large mammals or the lower survival and higher reproductive output that characterizes insects or amphibians). Higher detection probabilities alone did little to reduce bias in demographic rates. However, higher detection probabilities in count surveys were more effective at reducing uncertainty in abundance trends than using more of the available datasets.

Our results suggest that the unique context of a given monitoring program and study species will largely determine whether collecting multiple data streams versus collecting a single high-quality data stream (e.g., increasing detection probabilities or sample sizes) will most improve parameter estimates. Practitioners should carefully consider the best ways of answering the specific ecological or management questions of interest (e.g., estimating demographic rates versus detecting trends) given the species’ life history and the costs of data collection. This work provides a foundation from which to further explore the performance of IPMs and to inform the design of monitoring programs while considering the costs and benefits of data collection.

## Scripts, current version

### 0 - preparing scenarios

#### compute_time_calc.R
Code for approximating the total run time, given number of simulated datasets, number of parameter scenarios, computer cores, and estimated run time for each model.

#### generate_scenarios.R

Function 'getNviable' for finding the combinations of parameters that give a population growth rate ($\lambda$) within certain bounds, using the eigenvalue from the Leslie matrix. The output from this function produces the parameter scenarios used to simulate population trajectories, observation data from the trajectories, and fit the the IPM models.

### 1 - simulating data

#### IPM_sim_2.0function.R

Script contains two functions for simulating data: 'simPopTrajectory' and 'simData'.

The 'simPopTrajectory' function simulates the true population trajectory by individual and age class, starting from the stable age distribution. Individual fates are retained throughout total number of years.

The 'simData' function uses the output from the 'simPopTrajectory' function to simulate the observation process for each data type. The true individual data is broken down into independent sets, each is subject to observation error. Mark-resight, count, and reproductive success data are output from this function.

#### simulateTrajectories.R

Script that pulls from the scenarios that were created in '0 - preparing scenarios' and the functions in IPM_sim_2.0function.R to simulate observation data sets for model fitting.

### 2 - models

#### IPMinitvalues.R
Functions for putting data in the form required by the survival model.

#### IPM_marray.R
Script contains 4 IPM NIMBLE models.

 1. full IPM model ('IPMmod') where all three datasets are included
 2. model without productivity dataset ('nonests')
 3. model without survival model ('nomr')
 4. abundance only model without productivity or survival ('abundonly')

### 3 - run models

#### run_scenarios_helperFns.R

Script with functions and code to run NIMBLE models. Function 'marray' transforms the survival data capture histories to m-array format. MCMC settings in this script, with functions for each of the 4 models. Each model has a run model function that defines constants, initial values, data, parameters to monitor, and code to build/run the models in NIMBLE.

#### run_scenarios.R

Main workhorse script that pulls from the scripts above to run models in parallel and output results.

### 4 - process results

#### 01_load_and_process_data.R

Script that checks the Gelman-Rubin convergence statistic for each output MCMC sample, then thins the chains to reduce file size and places the output into categories based on the population growth rate, $\lambda$.

#### 02_compute_geom_means.R

Computes the geometric means for $\lambda$ in each MCMC output.

## Data

Contains true parameter scenarios used to simulate population trajectories and observation data, all viable parameter combinations in files '-.lam.combos.RDS' and the selected parameter combinations used to simulate data are in files '-.lam.params.RDS' *(where "-" is low, med, or high)*. Folders labeled '-Trajectories' contain the simulated population trajectories for each of the scenarios found in '-.lam.params.RDS'. Each simulated trajectory is labeled with the scenario and the simulation number for that scenario. For example, in the 'highTrajectory' folder: 'highpopTraj-1-2.RDS' corresponds to the first parameter scenario in 'high.lam.param.RDS' and is the second population trajectory simulated using those parameter values.

File 'true.vals.csv' is the true parameter values used in each scenario. 'scenario_ID.csv' tracks the combinations of detection probabilities and IPM models.